
Healthcare facilities, particularly in resource-constrained environments, face persistent challenges in medication delivery due to staff overload, missed doses, and human error. While service robots have shown promise in improving efficiency and reducing contact-based interactions, existing medical robotic systems are often expensive, non-modular, and lack intelligent, patient-specific functionality. This proposal presents a low-cost, modular autonomous medical assistant robot designed to address these limitations through intelligent automation and personalized care.

The proposed system employs line-following navigation using IR sensors, making it well suited for structured indoor environments such as hospital wards, hostels, and elderly care centers. The robot features a three-compartment modular architecture, consisting of a bottom mobility and sensor unit, a middle medical storage compartment, and a top AI-enabled smart pill dispensing module. This modular design allows for easy maintenance, scalability, and future upgrades.

A Raspberry Pi--based AI system with camera integration enables facial recognition for patient identification and time-scheduled pill dispensing, significantly reducing the risk of incorrect dosage and improving medication adherence. Additionally, wireless connectivity via Wi-Fi/Bluetooth supports real-time monitoring, schedule updates, and medication log tracking. By combining autonomous navigation, intelligent pill dispensing, and modular hardware design, the proposed solution enhances accuracy, reduces healthcare staff workload, and offers a scalable, cost-effective approach to smart healthcare automation.

