\begin{enumerate}
  \item X. Zhang, M. Liu, and Y. Zhao, “Design of angle detection system based on MPU6050,” in \textit{Proc. Int. Conf. Intelligent Transportation, Big Data \& Smart City}, Changsha, China, 2017, pp. 159–162.
  \item S. J. Park and S. Y. Lee, “Real-Time Motion Recognition Using an Accelerometer,” \textit{Sensors}, vol. 15, no. 1, pp. 135–148, 2015.
  \item P. Kumar, J. Verma, and S. Prasad, "Hand data glove: A wearable real-time device for human-computer interaction," International Journal of Advanced Science and Technology, vol. 43, pp. 15–26, June 2012. 
  \item A. Khandual and A. Biswal, “IoT Based Smart Glove for Hand Gesture Recognition,” \textit{Int. J. Sci. Res. Comput. Sci. Eng. Inf. Technol.}, vol. 5, no. 2, pp. 384–389, Mar. 2019.
  \item M. Nasri et al., “Design and Implementation of a 3D Printed Hand Gesture Recognition System Using Arduino and Flex Sensors,” \textit{IEEE Access}, vol. 7, pp. 55371–55378, 2019.
  \item T. Baudel and M. Beaudouin-Lafon, “Charade: Remote Control of Objects Using Free-Hand Gestures,” \textit{Commun. ACM}, vol. 36, no. 7, pp. 28–35, 1993.
  \item S. S. Rautaray and A. Agrawal, "Real time hand gesture recognition system for dynamic applications," International Journal of UbiComp (IJU), vol. 3, no. 1, pp. 21–35, Jan. 2012.
  \item S. Chowdhury and A. Haque, "Animatronic hand controller," Technical Report, Dept. of Electrical and Electronic Engineering, Ahsanullah University of Science and Technology, Dhaka, Bangladesh, Dec. 2013.
  \item D. J. Sturman and D. Zeltzer, "A survey of glove-based input," IEEE Computer Graphics and Applications, vol. 14, no. 1, pp. 30–39, Jan. 1994.
  \item J. Huang, "Design of angle detection system based on MPU6050," in Proc. 7th Int. Conf. Education, Management, Information and Computer Science (ICEMC 2017), Advances in Computer Science Research, vol. 73, pp. 1–3, 2017.
\end{enumerate}