The advancement of healthcare robotics has led to the development of autonomous hospital service robots, intelligent medicine delivery systems, and automated pill dispensing mechanisms. These systems aim to improve efficiency, reduce human error, and enhance patient care. Various studies have explored different aspects of healthcare automation; however, most of them focus on specific functionalities rather than providing a unified solution.

A line-following robot for hospital operations was developed in \cite{ref1}, where the system was designed to transport medicines, linens, and medical equipment along predefined paths within hospital environments. The robot utilized infrared (IR) sensors for path detection and an Arduino Mega microcontroller for navigation and control.The robot's movement is typically achieved using DC geared motors such as Johnson motors, which provide high torque at low speeds and are widely used in robotic applications for precise motion control \cite{ref8}. A Bluetooth-based Android interface enabled remote supervision. The study demonstrated that low-cost embedded systems can improve hospital logistics efficiency. However, it lacked intelligent decision-making, patient-specific identification, and automated dispensing capabilities, limiting its application in advanced healthcare systems. The effectiveness of IR sensors in such applications is further supported in \cite{ref7}, where IR sensors are described as reliable components for detecting surface variations based on reflected infrared radiation. These sensors use an IR LED and photodiode pair, making them suitable for line-following robots due to their low cost, fast response, and non-contact operation.

In another study \cite{ref2}, a Bluetooth-controlled medicine delivery robot named \textit{MedBuddy} was proposed to reduce direct interaction between healthcare workers and patients. The system incorporated ultrasonic sensors for obstacle detection and provided real-time video streaming. While it improved safety during infectious conditions, its reliance on manual control reduced autonomy, and the absence of intelligent scheduling and patient-specific dispensing limited its overall effectiveness.

An automated pill dispensing system using Raspberry Pi was presented in \cite{ref3}, where a servo-controlled mechanism was used along with a real-time clock (RTC) to dispense medicines at scheduled intervals. The system also implemented facial recognition for patient identification, enhancing medication adherence. However, the system was stationary and lacked mobility, which restricted its integration into broader hospital logistics.

Recent advancements in artificial intelligence have enabled the development of smart healthcare systems, particularly in patient identification and monitoring. Deep learning-based face recognition techniques have shown high accuracy in controlled environments, allowing personalized healthcare services. However, deploying such models on resource-constrained platforms like Raspberry Pi remains challenging due to computational limitations.

Similarly, IoT-based healthcare systems provide real-time monitoring, remote access, and data-driven decision-making by integrating sensors, cloud platforms, and web applications. Despite these advantages, many existing IoT solutions focus primarily on monitoring rather than actuation, lacking automated physical interaction such as medicine dispensing.

\begin{table}[H]
\centering
\begin{tabular}{|p{4.5cm}|p{5cm}|p{4.5cm}|}
\hline
\textbf{Feature} & \textbf{Existing Systems} & \textbf{Proposed System} \\
\hline
Autonomous Navigation & Limited & Yes \\
Face Recognition & Partial / None & Yes \\
Automated Dispensing & Basic & Advanced \\
Real-time Monitoring & Limited & Yes \\
Mobility & No / Partial & Yes \\
Cost & High & Low \\
\hline
\end{tabular}
\caption{Comparison Between Existing Systems and Proposed Medicare Bot}
\label{tab:system_comparison}
\end{table}

From the reviewed literature, it is evident that most existing systems address individual aspects of healthcare automation such as navigation, dispensing, or monitoring independently. There is a lack of an integrated system that combines autonomous navigation, AI-based patient identification, real-time monitoring, and automated medication dispensing within a single platform. Additionally, many solutions are either costly or not scalable for deployment in resource-constrained environments. Therefore, there is a clear research gap in developing a unified, low-cost, and intelligent healthcare system that integrates these functionalities into a cohesive and efficient solution.