
\subsection{Overview}

Although the Medicare Bot system demonstrates successful integration of AI, embedded systems, and web technologies, certain limitations exist due to hardware constraints, environmental factors, and system design decisions.

\subsection{Hardware Limitations}

The system is built using a Raspberry Pi 3B, which has limited processing power and memory. This restricts the use of more complex and computationally intensive deep learning models.

Additionally, servo motors require stable power supply. Any fluctuation in voltage can affect dispensing accuracy and reliability.

\subsection{Face Recognition Limitations}

The face recognition system is sensitive to environmental conditions. Performance may degrade under:

\begin{itemize}
    \item Poor lighting conditions
    \item Partial occlusions such as masks or glasses
    \item Extreme facial angles
\end{itemize}

The system currently supports recognition of one person at a time. If multiple faces are detected, the system blocks dispensing to avoid incorrect identification.

\subsection{System Logic Limitations}

The medication dispensing logic depends on predefined schedules. Currently, the system lacks advanced features such as:

\begin{itemize}
    \item Automatic detection of missed doses
    \item Adaptive scheduling based on patient behavior
\end{itemize}

Additionally, improper handling of the \texttt{last\_dispensed} field can lead to repeated dispensing if not correctly updated.

\subsection{Network Dependency}

The system relies on network connectivity for communication between the Raspberry Pi and the backend server. In case of network failure:

\begin{itemize}
    \item Logs may not be transmitted
    \item Real-time monitoring may be interrupted
\end{itemize}

\subsection{Scalability Limitations}

The current implementation uses SQLite, which is suitable for small-scale systems but may not perform efficiently for large-scale deployments with multiple users and devices.

\subsection{Security Limitations}

Although API key authentication is implemented, the system does not include advanced security features such as:

\begin{itemize}
    \item Encrypted communication (HTTPS)
    \item Role-based access control
    \item Biometric spoof detection
\end{itemize}

\subsection{Mechanical Limitations}

The dispensing mechanism is based on servo motors and fixed compartments. This design:

\begin{itemize}
    \item Limits the number of medications that can be stored
    \item Requires manual refilling
    \item May face wear and tear over long-term use
\end{itemize}